---
title: "INFM600_0201_TeamExceptional_RScript"
output: html_notebook
---

INFM600 Information Environment[Team Exceptional]
Document: R Script
Date: 10 December 2016

Document Structure: We have three research questions. This document contains three separate scripts for each question.

Research Question:[1] Does the change in temperature have an effect on the crime rate?
Description: We are analyzing the trend of crime rate with respect to the changes in the temperature. Simple linear regression model is applied to figure out whether there is a relationship between the crime rate and the temperature.
Dataset/s: Weather_Crime_Incidents2011-15 
NOTE: This dataset includes the details of the crimes as well as the weather details of that day. This was created from two separate datasets in the data-cleaning phase
Columns/Attributes used: REPORTDATENAME, OFFENSE,START_DATE,MaxTemp,MinTemp,RangeTemp
Packages: plyr, ggplot2


```{r RQ1 Part 1}
# [Step 0] Reading the dataset into RStudio
data= read.csv(file.choose())

#Load support libraries
# Load rename function
library(plyr)
# Load plot function
library(ggplot2)
# Load xkcd 
library(xkcd)
library(extrafont)

# XKCD theme
theme_xkcd <- theme(
  panel.background = element_rect(fill="white"),
  panel.grid = element_line(colour="white"),
  axis.text.x = element_text(colour="black"),
  text = element_text(size=8)
)

#Part 1: Check the relationship between the frequency of all cases and temperature (Max Temp and Min Temp).
# [Step 1] Calculating the mean and standard deviation of maximum temperature and minimum temperature
mean(data$MaxTemp)
sd(data$MaxTemp)
mean(data$MinTemp)
sd(data$MinTemp)
#  [Step 2] Draw line chart
#Mean and standard deviation of max temperature
mean(data$MaxTemp)
sd(data$MaxTemp)
#Mean and standard deviation of min temperature
mean(data$MinTemp)
sd(data$MinTemp)
#Min Temperature
minT=table(data$MinTemp)
minT=as.data.frame(minT)
minT$Var1=as.numeric(as.character(minT$Var1))
minT=rename(minT, c("Var1"="Min.Temp", "Freq"="NumberofCases"))
#Max Temperature
maxT=table(data$MaxTemp)
maxT=as.data.frame(maxT)
maxT$Var1=as.numeric(as.character(maxT$Var1))
maxT=rename(maxT, c("Var1"="Max.Temp", "Freq"="NumberofCases"))

# Draw scatter plot
plot = ggplot() +
  geom_point(data=maxT,aes(x=Max.Temp,y=NumberofCases),color = "red")+
  geom_point(data=minT,aes(x=Min.Temp,y=NumberofCases),color = "blue")+
  labs(title = "The distribution of cases happened from 2011-2015", x="Temperature in degrees Fahrenheit", y="Number of cases")+ 
  theme_xkcd+
  theme(plot.title = element_text(hjust = 0.5))

plot
# [Step 3] Classifying the data of 2011-2015 according to the type of crimes and counting the frequency for each type
#ARSON
arson=subset(data,OFFENSE=="ARSON")
nrow(arson)

#ASSAULT
assault=subset(data,OFFENSE=="ASSAULT W/DANGEROUS WEAPON")
nrow(assault)

#BURGLARY
burglary=subset(data,OFFENSE=="BURGLARY")
nrow(burglary)

#HOMICIDE
homicide=subset(data,OFFENSE=="HOMICIDE")
nrow(homicide)

#MOTOR VEHICLE THEFT
mvt=subset(data,OFFENSE=="MOTOR VEHICLE THEFT")
nrow(mvt)

#ROBBERY
robbery=subset(data,OFFENSE=="ROBBERY")
nrow(robbery)

#SEX ABUSE
sa=subset(data,OFFENSE=="SEX ABUSE")
nrow(sa)

#THEFT FROM AUTO
thefa=subset(data,OFFENSE=="THEFT F/AUTO")
nrow(thefa)

#THEFT FROM OTHER
thefot=subset(data,OFFENSE=="THEFT/OTHER")
nrow(thefot)
```



Part 2: Compare the impact of the range of temperature on the occurrence rate of all types of cases.
Independent variable: Range of temperature of each day (interval)
Dependent varibale: The frequency of each type of cases happened (ratio)

```{r RQ1 Part 2}
# [Step 1] Calculating the frequency of each type of cases from the data of 2011-2015 corresponding to each numerical value in "range of temperature" column.

#ARSON
rangeT_arson=table(arson$RangeTemp)
rangeT_arson=as.data.frame(rangeT_arson)
rangeT_arson$Var1=as.numeric(as.character(rangeT_arson$Var1))
rangeT_arson=rename(rangeT_arson, c("Var1"="Range.of.Temp", "Freq"="NumberofCases"))

#ASSAULT
rangeT_assault=table(assault$RangeTemp)
rangeT_assault=as.data.frame(rangeT_assault)
rangeT_assault$Var1=as.numeric(as.character(rangeT_assault$Var1))
rangeT_assault=rename(rangeT_assault, c("Var1"="Range.of.Temp", "Freq"="NumberofCases"))

#BURGLARY
rangeT_burglary=table(burglary$RangeTemp)
rangeT_burglary=as.data.frame(rangeT_burglary)
rangeT_burglary$Var1=as.numeric(as.character(rangeT_burglary$Var1))
rangeT_burglary=rename(rangeT_burglary, c("Var1"="Range.of.Temp", "Freq"="NumberofCases"))

#HOMICIDE
rangeT_homicide=table(homicide$RangeTemp)
rangeT_homicide=as.data.frame(rangeT_homicide)
rangeT_homicide$Var1=as.numeric(as.character(rangeT_homicide$Var1))
rangeT_homicide=rename(rangeT_homicide, c("Var1"="Range.of.Temp", "Freq"="NumberofCases"))

#MOTOR VEHICLE THEFT
rangeT_mvt=table(mvt$RangeTemp)
rangeT_mvt=as.data.frame(rangeT_mvt)
rangeT_mvt$Var1=as.numeric(as.character(rangeT_mvt$Var1))
rangeT_mvt=rename(rangeT_mvt, c("Var1"="Range.of.Temp", "Freq"="NumberofCases"))

#ROBBERY
rangeT_robbery=table(robbery$RangeTemp)
rangeT_robbery=as.data.frame(rangeT_robbery)
rangeT_robbery$Var1=as.numeric(as.character(rangeT_robbery$Var1))
rangeT_robbery=rename(rangeT_robbery, c("Var1"="Range.of.Temp", "Freq"="NumberofCases"))

#SEX ABUSE
rangeT_sa=table(sa$RangeTemp)
rangeT_sa=as.data.frame(rangeT_sa)
rangeT_sa$Var1=as.numeric(as.character(rangeT_sa$Var1))
rangeT_sa=rename(rangeT_sa, c("Var1"="Range.of.Temp", "Freq"="NumberofCases"))

#THEFT FROM AUTO
rangeT_thefa=table(thefa$RangeTemp)
rangeT_thefa=as.data.frame(rangeT_thefa)
rangeT_thefa$Var1=as.numeric(as.character(rangeT_thefa$Var1))
rangeT_thefa=rename(rangeT_thefa, c("Var1"="Range.of.Temp", "Freq"="NumberofCases"))

#THEFT FROM OTHER
rangeT_thefot=table(thefot$RangeTemp)
rangeT_thefot=as.data.frame(rangeT_thefot)
rangeT_thefot$Var1=as.numeric(as.character(rangeT_thefot$Var1))
rangeT_thefot=rename(rangeT_thefot, c("Var1"="Range.of.Temp", "Freq"="NumberofCases"))

# [Step 2] Drawing histogram for each type
#ARSON
ggplot(data=rangeT_arson, aes(x=Range.of.Temp, y=NumberofCases, fill=Range.of.Temp)) +
  geom_bar(stat="identity")+
  labs(title = "The distribution of 'ARSON' cases  2011-2015", x="Range of temperature in degrees Fahrenheit", y="Number of cases")+
  theme_xkcd+
  theme(plot.title = element_text(hjust = 0.5))
#ASSAULT
ggplot(data=rangeT_assault, aes(x=Range.of.Temp, y=NumberofCases, fill=Range.of.Temp)) +
  geom_bar(stat="identity")+
  labs(title = "The distribution of 'ASSAULT' cases  2011-2015", x="Range of temperature in degrees Fahrenheit", y="Number of cases")+
  theme_xkcd+
  theme(plot.title = element_text(hjust = 0.5))
#BURGLARY
ggplot(data=rangeT_burglary, aes(x=Range.of.Temp, y=NumberofCases, fill=Range.of.Temp)) +
  geom_bar(stat="identity")+
  labs(title = "The distribution of 'BURGLARY' cases  2011-2015", x="Range of temperature in degrees Fahrenheit", y="Number of cases")+
  theme_xkcd+
  theme(plot.title = element_text(hjust = 0.5))
#HOMICIDE
ggplot(data=rangeT_homicide, aes(x=Range.of.Temp, y=NumberofCases, fill=Range.of.Temp)) +
  geom_bar(stat="identity")+
  labs(title = "The distribution of 'HOMICIDE' cases  2011-2015", x="Range of temperature in degrees Fahrenheit", y="Number of cases")+
  theme_xkcd+
  theme(plot.title = element_text(hjust = 0.5))
#MOTOR VEHICLE THEFT
ggplot(data=rangeT_mvt, aes(x=Range.of.Temp, y=NumberofCases, fill=Range.of.Temp)) +
  geom_bar(stat="identity")+
  labs(title = "The distribution of 'MOTOR VEHICLE THEFT' cases  2011-2015", x="Range of temperature in degrees Fahrenheit", y="Number of cases")+
  theme_xkcd+
  theme(plot.title = element_text(hjust = 0.5))
#ROBBERY
ggplot(data=rangeT_robbery, aes(x=Range.of.Temp, y=NumberofCases, fill=Range.of.Temp)) +
  geom_bar(stat="identity")+
  labs(title = "The distribution of 'ROBBERY' cases  2011-2015", x="Range of temperature in degrees Fahrenheit", y="Number of cases")+
  theme_xkcd+
  theme(plot.title = element_text(hjust = 0.5))
#SEX ABUSE
ggplot(data=rangeT_sa, aes(x=Range.of.Temp, y=NumberofCases, fill=Range.of.Temp)) +
  geom_bar(stat="identity")+
  labs(title = "The distribution of 'SEX ABUSE' cases  2011-2015", x="Range of temperature in degrees Fahrenheit", y="Number of cases")+
  theme_xkcd+
  theme(plot.title = element_text(hjust = 0.5))
#THEFT FROM AUTO
ggplot(data=rangeT_thefa, aes(x=Range.of.Temp, y=NumberofCases, fill=Range.of.Temp)) +
  geom_bar(stat="identity")+
  labs(title = "The distribution of 'THEFT FROM AUTO' cases  2011-2015", x="Range of temperature in degrees Fahrenheit", y="Number of cases")+
  theme_xkcd+
  theme(plot.title = element_text(hjust = 0.5))
#THEFT FROM OTHER
ggplot(data=rangeT_thefot, aes(x=Range.of.Temp, y=NumberofCases, fill=Range.of.Temp)) +
  geom_bar(stat="identity")+
  labs(title = "The distribution of 'THEFT FROM OTHER' cases 2011-2015", x="Range of temperature in degrees Fahrenheit", y="Number of cases")+
  theme_xkcd+
  theme(plot.title = element_text(hjust = 0.5))

```


Part 3: Compare the impact of max temperature on the occurrence rate of all types of cases.
Find the frequency of each type of cases from the data of 2011-2015 corresponding to each numerical value in "max temperature" column.
Independent varibale: Max Temprature of each day from 2011 to 2015 (interval)
Dependent varibale: The frequency of each type of cases happened (ratio)

```{r RQ1 PArt 3}
#ARSON
# [Step 1] Finding out the max temperature on the day when this type of cases happened, and then counting the frequency of the same type of cases corresponding to each temperature.
maxT_arson=table(arson$MaxTemp)
maxT_arson=as.data.frame(maxT_arson)
# [Step 2] Saving the value of temperature and  frequency of the type of cases in a new data frame
maxT_arson$Var1=as.numeric(as.character(maxT_arson$Var1))
# [Step 3] Renaming columns' names
maxT_arson=rename(maxT_arson, c("Var1"="MaxTemp", "Freq"="NumberofCases"))

#ASSAULT
# [Step 1] Finding out the max temperature on the day when this type of cases happened, and then counting the frequency of the same type of cases corresponding to each temperature.
maxT_assault=table(assault$MaxTemp)
maxT_assault=as.data.frame(maxT_assault)
# [Step 2] Saving the value of temperature and  frequency of the type of cases in a new data frame
maxT_assault$Var1=as.numeric(as.character(maxT_assault$Var1))
# [Step 3] Renaming columns' names
maxT_assault=rename(maxT_assault, c("Var1"="MaxTemp", "Freq"="NumberofCases"))

#BURGLARY
# [Step 1] Finding out the max temperature on the day when this type of cases happened, and then counting the frequency of the same type of cases corresponding to each temperature.
maxT_burglary=table(burglary$MaxTemp)
maxT_burglary=as.data.frame(maxT_burglary)
# [Step 2] Saving the value of temperature and  frequency of the type of cases in a new data frame
maxT_burglary$Var1=as.numeric(as.character(maxT_burglary$Var1))
# [Step 3] Renaming columns' names
maxT_burglary=rename(maxT_burglary, c("Var1"="MaxTemp", "Freq"="NumberofCases"))

#HOMICIDE
# [Step 1] Finding out the max temperature on the day when this type of cases happened, and then counting the frequency of the same type of cases corresponding to each temperature.
maxT_homicide=table(homicide$MaxTemp)
maxT_homicide=as.data.frame(maxT_homicide)
# [Step 2] Saving the value of temperature and  frequency of the type of cases in a new data frame
maxT_homicide$Var1=as.numeric(as.character(maxT_homicide$Var1))
# [Step 3] Renaming columns' names
maxT_homicide=rename(maxT_homicide, c("Var1"="MaxTemp", "Freq"="NumberofCases"))

#MOTOR VEHICLE THEFT
# [Step 1] Finding out the max temperature on the day when this type of cases happened, and then counting the frequency of the same type of cases corresponding to each temperature.
maxT_mvt=table(mvt$MaxTemp)
maxT_mvt=as.data.frame(maxT_mvt)
# [Step 2] Saving the value of temperature and  frequency of the type of cases in a new data frame
maxT_mvt$Var1=as.numeric(as.character(maxT_mvt$Var1))
# [Step 3] Renaming columns' names
maxT_mvt=rename(maxT_mvt, c("Var1"="MaxTemp", "Freq"="NumberofCases"))

#ROBBERY
# [Step 1] Finding out the max temperature on the day when this type of cases happened, and then counting the frequency of the same type of cases corresponding to each temperature.
maxT_robbery=table(robbery$MaxTemp)
maxT_robbery=as.data.frame(maxT_robbery)
# [Step 2] Saving the value of temperature and  frequency of the type of cases in a new data frame
maxT_robbery$Var1=as.numeric(as.character(maxT_robbery$Var1))
# [Step 3] Renaming columns' names
maxT_robbery=rename(maxT_robbery, c("Var1"="MaxTemp", "Freq"="NumberofCases"))

#SEX ABUSE
# [Step 1] Finding out the max temperature on the day when this type of cases happened, and then counting the frequency of the same type of cases corresponding to each temperature.
maxT_sa=table(sa$MaxTemp)
maxT_sa=as.data.frame(maxT_sa)
# [Step 2] Saving the value of temperature and  frequency of the type of cases in a new data frame
maxT_sa$Var1=as.numeric(as.character(maxT_sa$Var1))
# [Step 3] Renaming columns' names
maxT_sa=rename(maxT_sa, c("Var1"="MaxTemp", "Freq"="NumberofCases"))

#THEFT FROM AUTO
# [Step 1] Finding out the max temperature on the day when this type of cases happened, and then counting the frequency of the same type of cases corresponding to each temperature.
maxT_thefa=table(thefa$MaxTemp)
maxT_thefa=as.data.frame(maxT_thefa)
# [Step 2] Saving the value of temperature and  frequency of the type of cases in a new data frame
maxT_thefa$Var1=as.numeric(as.character(maxT_thefa$Var1))
# [Step 3] Renaming columns' names
maxT_thefa=rename(maxT_thefa, c("Var1"="MaxTemp", "Freq"="NumberofCases"))

#THEFT FROM OTHER
# [Step 1] Finding out the max temperature on the day when this type of cases happened, and then counting the frequency of the same type of cases corresponding to each temperature.
maxT_thefot=table(thefot$MaxTemp)
maxT_thefot=as.data.frame(maxT_thefot)
# [Step 2] Saving the value of temperature and  frequency of the type of cases in a new data frame
maxT_thefot$Var1=as.numeric(as.character(maxT_thefot$Var1))
# [Step 3] Renaming columns' names
maxT_thefot=rename(maxT_thefot, c("Var1"="MaxTemp", "Freq"="NumberofCases"))

# Part 4: Using "simple linear regression model" to analyze the relationship between IV and DV,because IV is interval,and DV is ratio.

#ARSON
summary(lm(NumberofCases~MaxTemp,data=maxT_arson))
#ASSAULT
summary(lm(NumberofCases~MaxTemp,data=maxT_assault))
#BURGLARY
summary(lm(NumberofCases~MaxTemp,data=maxT_burglary))
#HOMICIDE
summary(lm(NumberofCases~MaxTemp,data=maxT_homicide))
#MOTOR VEHICLE THEFT
summary(lm(NumberofCases~MaxTemp,data=maxT_mvt))
#ROBBERY
summary(lm(NumberofCases~MaxTemp,data=maxT_robbery))
#SEX ABUSE
summary(lm(NumberofCases~MaxTemp,data=maxT_sa))
#THEFT FROM AUTO
summary(lm(NumberofCases~MaxTemp,data=maxT_thefa))
#THEFT FROM OTHER
summary(lm(NumberofCases~MaxTemp,data=maxT_thefot))

# Part 5: Creating scatterplots

#ARSON
ggplot(maxT_arson,aes(MaxTemp,NumberofCases))+ 
  geom_point()+
  geom_smooth(method='lm')+
  labs(title = "The relationship between number of 'ARSON' cases and max temperature 2011-2015", x="Temperature in degrees Fahrenheit", y="Number of cases")+
  theme_xkcd+
  theme(plot.title = element_text(hjust = 0.5))
#ASSAULT
ggplot(maxT_assault,aes(MaxTemp,NumberofCases))+ 
  geom_point()+
  geom_smooth(method='lm')+
  labs(title = "The relationship between number of 'ASSAULT' cases and max temperature 2011-2015", x="Temperature in degrees Fahrenheit", y="Number of cases")+
  theme_xkcd+
  theme(plot.title = element_text(hjust = 0.5))
#BURGLARY
ggplot(maxT_burglary,aes(MaxTemp,NumberofCases))+ 
  geom_point()+
  geom_smooth(method='lm')+
  labs(title = "The relationship between number of 'BURGLARY' cases and max temperature 2011-2015", x="Temperature in degrees Fahrenheit", y="Number of cases")+
  theme_xkcd+
  theme(plot.title = element_text(hjust = 0.5))
#HOMICIDE
ggplot(maxT_homicide,aes(MaxTemp,NumberofCases))+ 
  geom_point()+
  geom_smooth(method='lm')+
  labs(title = "The relationship between number of 'HOMICIDE' cases and max temperature 2011-2015", x="Temperature in degrees Fahrenheit", y="Number of cases")+
  theme_xkcd+
  theme(plot.title = element_text(hjust = 0.5))
#MOTOR VEHICLE THEFT
ggplot(maxT_mvt,aes(MaxTemp,NumberofCases))+ 
  geom_point()+
  geom_smooth(method='lm')+
  labs(title = "The relationship between number of 'MOTOR VEHICLE THEFT' cases and max temperature 2011-2015", x="Temperature in degrees Fahrenheit", y="Number of cases")+
  theme_xkcd+
  theme(plot.title = element_text(hjust = 0.5))
#ROBBERY
ggplot(maxT_robbery,aes(MaxTemp,NumberofCases))+ 
  geom_point()+
  geom_smooth(method='lm')+
  labs(title = "The relationship between number of 'ROBBERY' cases and max temperature 2011-2015", x="Temperature in degrees Fahrenheit", y="Number of cases")+
  theme_xkcd+
  theme(plot.title = element_text(hjust = 0.5))
#SEX ABUSE
ggplot(maxT_sa,aes(MaxTemp,NumberofCases))+ 
  geom_point()+
  geom_smooth(method='lm')+
  labs(title = "The relationship between number of 'SEX ABUSE' cases and max temperature 2011-2015", x="Temperature in degrees Fahrenheit", y="Number of cases")+
  theme_xkcd+
  theme(plot.title = element_text(hjust = 0.5))
#THEFT FROM AUTO
ggplot(maxT_thefa,aes(MaxTemp,NumberofCases))+ 
  geom_point()+
  geom_smooth(method='lm')+
  labs(title = "The relationship between number of 'THEFT FROM AUTO' cases and max temperature 2011-2015", x="Temperature in degrees Fahrenheit", y="Number of cases")+
  theme_xkcd+
  theme(plot.title = element_text(hjust = 0.5))
#THEFT FROM OTHER
ggplot(maxT_thefot,aes(MaxTemp,NumberofCases))+ 
  geom_point()+
  geom_smooth(method='lm')+
  labs(title = "The relationship between  number of 'THEFT FROM OTHER' cases and max temperature 2011-2015", x="Temperature in degrees Fahrenheit", y="Number of cases")+
  theme_xkcd+
  theme(plot.title = element_text(hjust = 0.5))
# End of R Script for Research Question [1]
```

Research Question:[2]	Is there a relation between the number of nightclubs with the crime rate in a particular geographic unit (cluster)?
Description: We are analyzing the crime rate(number of crimes) for each cluster with respect to the number of nightclubs in each cluster. Moreover, we will be looking into different types of crimes (Homicide, Burglary, Sexual Assault, etc.) We will be running the correlation test for each type of crime to figure out if there is a relationship between the number of nightclubs and occurence of crime
Datasets: (1) Crime_Incidents_year(2011-2015)- all details about crimes for each year
    			(2) NightClub_Cluster- list of nightclubs classified by cluster id
NOTE: The above mentioned datasets are the ones we created after data-cleaning phase.
Columns/Attributes used: NEIGHBORHOODCLUSTER, OFFENCE in dataset(1), NEIGHBORHOODCLUSTER, NUMBER_CLUBS in dataset(2)
Packages: No extra packages are used

```{r RQ2 Part 1:Descriptive Analysis}
# [Step 0.0] Opening each dataset in RStudio; CrimeRate2011-15.csv and NightClub_Cluster.csv respectively
merged.dataset <- read.csv(file.choose())

club.clusterwise <- read.csv(file.choose())


# [Step 0.1] Merging the crime data for all years 2011-2015 into one dataset

#merged.dataset=rbind(Crime_Incidents__2015,Crime_Incidents__2012)
#merged.dataset=rbind(merged.dataset,Crime_Incidents__2013)
#merged.dataset=rbind(merged.dataset,Crime_Incidents__2014)
#merged.dataset=rbind(merged.dataset,Crime_Incidents__2015)

# Crime Type: ALL
# [Step 1] Creating a frequency table to get the number of crimes for each cluster and writing it to a csv file for recordkeeping
total.clusterwise=table(merged.dataset$NEIGHBORHOODCLUSTER)
write.csv(total.clusterwise,file="total.clusterwise.csv")

# [Step 2] Reading in the file created in above step and renaming the columns for better understandability 
total.clusterwise <- read.csv(file.choose())
#View(total.clusterwise)
names(total.clusterwise)=c("Cluster","Number.Crimes")

# [Step 3.0] Calculating the average number of crimes occuring in a cluster to be able to understand our data better
total.cluster.mean=mean(total.clusterwise$Number.Crimes)
total.cluster.mean
# [Step 3.1] Calculating the maximum and minimum number of crimes occuring in a cluster to figure out the cluster with the highest and lowest number of crimes
total.cluster.range=range(total.clusterwise$Number.Crimes)
total.cluster.range
# [Step 3.2] Calculating the standard deviation of crimes occuring in a cluster to get an idea about the variance seen in our data
total.cluster.sd=sd(total.clusterwise$Number.Crimes)
total.cluster.sd

# Crime Type: ARSON
# [Step 1] Creating a subset of those rows which describe crimes of type:ARSON 
arson=subset(merged.dataset,merged.dataset$OFFENSE=="ARSON")

# [Step 2] Creating a frequency table to get the number of crimes for each cluster and writing it to a csv file for recordkeeping
arson.clusterwise=table(arson$NEIGHBORHOODCLUSTER)
write.csv(arson.clusterwise,file="arson.clusterwise.csv")

# [Step 3] Reading in the file created in above step and renaming the columns for better understandability 
arson.clusterwise <- read.csv(file.choose())

names(arson.clusterwise)=c("Cluster","Number.Crimes")

# [Step 4.0] Calculating the average number of crimes occuring in a cluster to be able to understand our data better
arson.cluster.mean=mean(arson.clusterwise$Number.Crimes)
arson.cluster.mean
# [Step 4.1] Calculating the maximum and minimum number of crimes occuring in a cluster to figure out the cluster with the highest and lowest number of crimes
arson.cluster.range=range(arson.clusterwise$Number.Crimes)
arson.cluster.range
# [Step 4.2] Calculating the standard deviation of crimes occuring in a cluster to get an idea about the variance seen in our data
arson.cluster.sd=sd(arson.clusterwise$Number.Crimes)
arson.cluster.sd

# Crime Type: ASSAULT W/DANGEROUS WEAPON
# [Step 1] Creating a subset of those rows which describe crimes of type:ASSAULT W/DANGEROUS WEAPON 
assault=subset(merged.dataset,merged.dataset$OFFENSE=="ASSAULT W/DANGEROUS WEAPON")

# [Step 2] Creating a frequency table to get the number of crimes for each cluster and writing it to a csv file for recordkeeping
assault.clusterwise=table(assault$NEIGHBORHOODCLUSTER)
write.csv(assault.clusterwise,file="assault.clusterwise.csv")

# [Step 3] Reading in the file created in above step and renaming the columns for better understandability 
assault.clusterwise <- read.csv(file.choose())
#View(assault.clusterwise)
names(assault.clusterwise)=c("Cluster","Number.Crimes")

# [Step 4.0] Calculating the average number of crimes occuring in a cluster to be able to understand our data better
#assault.cluster.mean=mean(assault.clusterwise$Number.Crimes)
assault.cluster.mean
# [Step 4.1] Calculating the maximum and minimum number of crimes occuring in a cluster to figure out the cluster with the highest and lowest number of crimes
assault.cluster.range=range(assault.clusterwise$Number.Crimes)
assault.cluster.range
# [Step 4.2] Calculating the standard deviation of crimes occuring in a cluster to get an idea about the variance seen in our data
assault.cluster.sd=sd(assault.clusterwise$Number.Crimes)
assault.cluster.sd

# Crime Type: BURGLARY
# [Step 1] Creating a subset of those rows which describe crimes of type:BURGLARY 
burglary=subset(merged.dataset,merged.dataset$OFFENSE=="BURGLARY")

# [Step 2] Creating a frequency table to get the number of crimes for each cluster and writing it to a csv file for recordkeeping
burglary.clusterwise=table(burglary$NEIGHBORHOODCLUSTER)
write.csv(burglary.clusterwise,file="burglary.clusterwise.csv")

# [Step 3] Reading in the file created in above step and renaming the columns for better understandability 
burglary.clusterwise <- read.csv(file.choose())
#View(burglary.clusterwise)
names(burglary.clusterwise)=c("Cluster","Number.Crimes")

# [Step 4.0] Calculating the average number of crimes occuring in a cluster to be able to understand our data better
burglary.cluster.mean=mean(burglary.clusterwise$Number.Crimes)
burglary.cluster.mean
# [Step 4.1] Calculating the maximum and minimum number of crimes occuring in a cluster to figure out the cluster with the highest and lowest number of crimes
burglary.cluster.range=range(burglary.clusterwise$Number.Crimes)
burglary.cluster.range
# [Step 4.2] Calculating the standard deviation of crimes occuring in a cluster to get an idea about the variance seen in our data
burglary.cluster.sd=sd(burglary.clusterwise$Number.Crimes)
burglary.cluster.sd

# Crime Type: HOMICIDE
# [Step 1] Creating a subset of those rows which describe crimes of type:HOMICIDE 
homicide=subset(merged.dataset,merged.dataset$OFFENSE=="HOMICIDE")

# [Step 2] Creating a frequency table to get the number of crimes for each cluster and writing it to a csv file for recordkeeping
homicide.clusterwise=table(homicide$NEIGHBORHOODCLUSTER)
write.csv(homicide.clusterwise,file="homicide.clusterwise.csv")

# [Step 3] Reading in the file created in above step and renaming the columns for better understandability 
homicide.clusterwise <- read.csv(file.choose())
#View(homicide.clusterwise)
names(homicide.clusterwise)=c("Cluster","Number.Crimes")

# [Step 4.0] Calculating the average number of crimes occuring in a cluster to be able to understand our data better
homicide.cluster.mean=mean(homicide.clusterwise$Number.Crimes)
homicide.cluster.mean
# [Step 4.1] Calculating the maximum and minimum number of crimes occuring in a cluster to figure out the cluster with the highest and lowest number of crimes
homicide.cluster.range=range(homicide.clusterwise$Number.Crimes)
homicide.cluster.range
# [Step 4.2] Calculating the standard deviation of crimes occuring in a cluster to get an idea about the variance seen in our data
homicide.cluster.sd=sd(homicide.clusterwise$Number.Crimes)
homicide.cluster.sd

# Crime Type: MOTOR VEHICLE THEFT
# [Step 1] Creating a subset of those rows which describe crimes of type:MOTOR VEHICLE THEFT 
motor.theft=subset(merged.dataset,merged.dataset$OFFENSE=="MOTOR VEHICLE THEFT")

# [Step 2] Creating a frequency table to get the number of crimes for each cluster and writing it to a csv file for recordkeeping
motor.theft.clusterwise=table(motor.theft$NEIGHBORHOODCLUSTER)
write.csv(motor.theft.clusterwise,file="motor.theft.clusterwise.csv")

# [Step 3] Reading in the file created in above step and renaming the columns for better understandability 
motor.theft.clusterwise <- read.csv(file.choose())
#View(motor.theft.clusterwise)
names(motor.theft.clusterwise)=c("Cluster","Number.Crimes")

# [Step 4.0] Calculating the average number of crimes occuring in a cluster to be able to understand our data better
motor.theft.cluster.mean=mean(motor.theft.clusterwise$Number.Crimes)
motor.theft.cluster.mean
# [Step 4.1] Calculating the maximum and minimum number of crimes occuring in a cluster to figure out the cluster with the highest and lowest number of crimes
motor.theft.cluster.range=range(motor.theft.clusterwise$Number.Crimes)
motor.theft.cluster.range
# [Step 4.2] Calculating the standard deviation of crimes occuring in a cluster to get an idea about the variance seen in our data
motor.theft.cluster.sd=sd(motor.theft.clusterwise$Number.Crimes)
motor.theft.cluster.sd

# Crime Type: ROBBERY
# [Step 1] Creating a subset of those rows which describe crimes of type:ROBBERY 
robbery=subset(merged.dataset,merged.dataset$OFFENSE=="ROBBERY")

# [Step 2] Creating a frequency table to get the number of crimes for each cluster and writing it to a csv file for recordkeeping
robbery.clusterwise=table(robbery$NEIGHBORHOODCLUSTER)
write.csv(robbery.clusterwise,file="robbery.clusterwise.csv")

# [Step 3] Reading in the file created in above step and renaming the columns for better understandability 
robbery.clusterwise <- read.csv(file.choose())
#View(robbery.clusterwise)
names(robbery.clusterwise)=c("Cluster","Number.Crimes")

# [Step 4.0] Calculating the average number of crimes occuring in a cluster to be able to understand our data better
robbery.cluster.mean=mean(robbery.clusterwise$Number.Crimes)
robbery.cluster.mean
# [Step 4.1] Calculating the maximum and minimum number of crimes occuring in a cluster to figure out the cluster with the highest and lowest number of crimes
robbery.cluster.range=range(robbery.clusterwise$Number.Crimes)
robbery.cluster.range
# [Step 4.2] Calculating the standard deviation of crimes occuring in a cluster to get an idea about the variance seen in our data
robbery.cluster.sd=sd(robbery.clusterwise$Number.Crimes)
robbery.cluster.sd

# Crime Type: SEX ABUSE
# [Step 1] Creating a subset of those rows which describe crimes of type:SEX ABUSE 
sex.abuse=subset(merged.dataset,merged.dataset$OFFENSE=="SEX ABUSE")

# [Step 2] Creating a frequency table to get the number of crimes for each cluster and writing it to a csv file for recordkeeping
sex.abuse.clusterwise=table(sex.abuse$NEIGHBORHOODCLUSTER)
write.csv(sex.abuse.clusterwise,file="sex.abuse.clusterwise.csv")

# [Step 3] Reading in the file created in above step and renaming the columns for better understandability 
sex.abuse.clusterwise <- read.csv(file.choose())
#View(sex.abuse.clusterwise)
names(sex.abuse.clusterwise)=c("Cluster","Number.Crimes")

# [Step 4.0] Calculating the average number of crimes occuring in a cluster to be able to understand our data better
sex.abuse.cluster.mean=mean(sex.abuse.clusterwise$Number.Crimes)
sex.abuse.cluster.mean
# [Step 4.1] Calculating the maximum and minimum number of crimes occuring in a cluster to figure out the cluster with the highest and lowest number of crimes
sex.abuse.cluster.range=range(sex.abuse.clusterwise$Number.Crimes)
sex.abuse.cluster.range
# [Step 4.2] Calculating the standard deviation of crimes occuring in a cluster to get an idea about the variance seen in our data
sex.abuse.cluster.sd=sd(sex.abuse.clusterwise$Number.Crimes)
sex.abuse.cluster.sd

# Crime Type: THEFT F/AUTO
# [Step 1] Creating a subset of those rows which describe crimes of type:THEFT F/AUTO 
theft.auto=subset(merged.dataset,merged.dataset$OFFENSE=="THEFT F/AUTO")

# [Step 2] Creating a frequency table to get the number of crimes for each cluster and writing it to a csv file for recordkeeping
theft.auto.clusterwise=table(theft.auto$NEIGHBORHOODCLUSTER)
write.csv(theft.auto.clusterwise,file="theft.auto.clusterwise.csv")

# [Step 3] Reading in the file created in above step and renaming the columns for better understandability 
theft.auto.clusterwise <- read.csv(file.choose())
#View(theft.auto.clusterwise)
names(theft.auto.clusterwise)=c("Cluster","Number.Crimes")

# [Step 4.0] Calculating the average number of crimes occuring in a cluster to be able to understand our data better
theft.auto.cluster.mean=mean(theft.auto.clusterwise$Number.Crimes)
theft.auto.cluster.mean
# [Step 4.1] Calculating the maximum and minimum number of crimes occuring in a cluster to figure out the cluster with the highest and lowest number of crimes
theft.auto.cluster.range=range(theft.auto.clusterwise$Number.Crimes)
theft.auto.cluster.range
# [Step 4.2] Calculating the standard deviation of crimes occuring in a cluster to get an idea about the variance seen in our data
theft.auto.cluster.sd=sd(theft.auto.clusterwise$Number.Crimes)
theft.auto.cluster.sd

# Crime Type: THEFT/OTHER
# [Step 1] Creating a subset of those rows which describe crimes of type:THEFT/OTHER 
theft.other=subset(merged.dataset,merged.dataset$OFFENSE=="THEFT/OTHER")

# [Step 2] Creating a frequency table to get the number of crimes for each cluster and writing it to a csv file for recordkeeping
theft.other.clusterwise=table(theft.other$NEIGHBORHOODCLUSTER)
write.csv(theft.other.clusterwise,file="theft.other.clusterwise.csv")

# [Step 3] Reading in the file created in above step and renaming the columns for better understandability 
theft.other.clusterwise <- read.csv(file.choose())
#View(theft.other.clusterwise)
names(theft.other.clusterwise)=c("Cluster","Number.Crimes")

# [Step 4.0] Calculating the average number of crimes occuring in a cluster to be able to understand our data better
theft.other.cluster.mean=mean(theft.other.clusterwise$Number.Crimes)
theft.other.cluster.mean
# [Step 4.1] Calculating the maximum and minimum number of crimes occuring in a cluster to figure out the cluster with the highest and lowest number of crimes
theft.other.cluster.range=range(theft.other.clusterwise$Number.Crimes)
theft.other.cluster.range
# [Step 4.2] Calculating the standard deviation of crimes occuring in a cluster to get an idea about the variance seen in our data
theft.other.cluster.sd=sd(theft.other.clusterwise$Number.Crimes)
theft.other.cluster.sd

#Clubs

# [Step 0] The relevant dataset is already open (see Step 0.0 at the beginning of the script). Here we rename the columns for better understandability
names(club.clusterwise)=c("Cluster","Number.Clubs")

# [Step 1.0] Calculating the average number of crimes occuring in a cluster to be able to understand our data better
club.cluster.mean=mean(club.clusterwise$Number.Clubs)
club.cluster.mean
# [Step 1.1] Calculating the maximum and minimum number of crimes occuring in a cluster to figure out the cluster with the highest and lowest number of crimes
club.cluster.range=range(club.clusterwise$Number.Clubs)
club.cluster.range
# [Step 1.2] Calculating the standard deviation of crimes occuring in a cluster to get an idea about the variance seen in our data
club.cluster.sd=sd(club.clusterwise$Number.Clubs)
club.cluster.sd
```

Correlation tests: For figuring out if there is a relationship between the number of nightclubs and number of crimes
```{r RQ2 PArt 2:Correlation Tests}
# Crime Type: ALL

# [Step 1] Merging the crime dataset with nightclubs dataset
club.total.cluster=merge(club.clusterwise,total.clusterwise,by="Cluster")

# [Step 2] Running the correlation test (linear regression model) between the number of clubs and number of crimes in a cluster
cor.test(club.total.cluster$Number.Clubs,club.total.cluster$Number.Crimes)


# Crime Type: ARSON

# [Step 1] Merging the crime dataset with nightclubs dataset
club.arson.cluster=merge(club.clusterwise,arson.clusterwise,by="Cluster")

# [Step 2] Running the correlation test (linear regression model) between the number of clubs and number of crimes in a cluster
cor.test(club.arson.cluster$Number.Clubs,club.arson.cluster$Number.Crimes)


# Crime Type: ASSAULT W/DANGEROUS WEAPON

# [Step 1] Merging the crime dataset with nightclubs dataset
club.assault.cluster=merge(club.clusterwise,assault.clusterwise,by="Cluster")

# [Step 2] Running the correlation test (linear regression model) between the number of clubs and number of crimes in a cluster
cor.test(club.assault.cluster$Number.Clubs,club.assault.cluster$Number.Crimes)


# Crime Type: BURGLARY

# [Step 1] Merging the crime dataset with nightclubs dataset
club.burglary.cluster=merge(club.clusterwise,burglary.clusterwise,by="Cluster")

# [Step 2] Running the correlation test (linear regression model) between the number of clubs and number of crimes in a cluster
cor.test(club.burglary.cluster$Number.Clubs,club.burglary.cluster$Number.Crimes)


# Crime Type: HOMICIDE

# [Step 1] Merging the crime dataset with nightclubs dataset
club.homicide.cluster=merge(club.clusterwise,homicide.clusterwise,by="Cluster")

# [Step 2] Running the correlation test (linear regression model) between the number of clubs and number of crimes in a cluster
cor.test(club.homicide.cluster$Number.Clubs,club.homicide.cluster$Number.Crimes)


# Crime Type: MOTOR VEHICLE THEFT

# [Step 1] Merging the crime dataset with nightclubs dataset
club.motor.theft.cluster=merge(club.clusterwise,motor.theft.clusterwise,by="Cluster")

# [Step 2] Running the correlation test (linear regression model) between the number of clubs and number of crimes in a cluster
cor.test(club.motor.theft.cluster$Number.Clubs,club.motor.theft.cluster$Number.Crimes)


# Crime Type: ROBBERY

# [Step 1] Merging the crime dataset with nightclubs dataset
club.robbery.cluster=merge(club.clusterwise,robbery.clusterwise,by="Cluster")

# [Step 2] Running the correlation test (linear regression model) between the number of clubs and number of crimes in a cluster
cor.test(club.robbery.cluster$Number.Clubs,club.robbery.cluster$Number.Crimes)


# Crime Type: SEX ABUSE

# [Step 1] Merging the crime dataset with nightclubs dataset
club.sex.abuse.cluster=merge(club.clusterwise,sex.abuse.clusterwise,by="Cluster")

# [Step 2] Running the correlation test (linear regression model) between the number of clubs and number of crimes in a cluster
cor.test(club.sex.abuse.cluster$Number.Clubs,club.sex.abuse.cluster$Number.Crimes)


# Crime Type: THEFT F/AUTO

# [Step 1] Merging the crime dataset with nightclubs dataset
club.theft.auto.cluster=merge(club.clusterwise,theft.auto.clusterwise,by="Cluster")

# [Step 2] Running the correlation test (linear regression model) between the number of clubs and number of crimes in a cluster
cor.test(club.theft.auto.cluster$Number.Clubs,club.theft.auto.cluster$Number.Crimes)


# Crime Type: THEFT/OTHER

# [Step 1] Merging the crime dataset with nightclubs dataset
club.theft.other.cluster=merge(club.clusterwise,theft.other.clusterwise,by="Cluster")

# [Step 2] Running the correlation test (linear regression model) between the number of clubs and number of crimes in a cluster
cor.test(club.theft.other.cluster$Number.Clubs,club.theft.other.cluster$Number.Crimes)

# End of RScript for Research Question [2]

```

Research Question:[3]	Is there a spike in the crime rate during major public holidays?
Description: We are analyzing the crime rate during the public holidays to check if there is a spike. We are further drilling down to specific types of crimes (eg: burgary, theft, etc) to figure out trends. Simple t-tests are carried out for basic analysis.
Datasets: (1) Crime_Incidents_year(2011-2015)- all details about crimes for each year
    			(2) Holidays- list of public holidays for United States of  America for the years (2011-2015	)
NOTE: The above mentioned datasets are the ones we created after data-cleaning phase.
Columns/Attributes used: REPORTDATETIME, START_DATE, END_DATE in dataset (1), Date in dataset(2)
Packages: dplyr, tidyr
```{r}
# [Step 1] Segregating the REPORTDATETIME column into Date and Time columns by using the library tidyr
library(tidyr)
FinalDataSet=merged.dataset
FinalDataSet <- separate(FinalDataSet, REPORTDATETIME, c("Date", "Time"), sep = "T")


#[Step 2] Converting Date column to proper format
#class(FinalDataSet$Date)
FinalDataSet$Date <- as.Date(FinalDataSet$Date)
#class(FinalDataSet$Date)

#[Step 3] Segregating the START_DATE AND END_DATE column, having timestamp into StartDate and EndDate columns without Timestamp
FinalDataSet <- separate(FinalDataSet, START_DATE , c("StartDate", "Time1"), sep = "T")
FinalDataSet <- separate(FinalDataSet, END_DATE, c("EndDate", "Time2"), sep = "T")


#[Step 4] Converting the StartDate and EndDate into Date format
FinalDataSet$StartDate <- as.Date(FinalDataSet$StartDate)
FinalDataSet$EndDate <- as.Date(FinalDataSet$EndDate)
```

After getting the required date, time and other necessary columns all the columns that were not required had to be deleted. A copy of the cleansed data was saved on the local system.
```{r Removing unnecessary columns and making a copy of cleansed data }
#[Step 5] removing all the unnecessary columns from the dataset
FinalDataSet$X <- NULL
FinalDataSet$Y <- NULL
FinalDataSet$OBJECTID <- NULL
FinalDataSet$CCN <- NULL
FinalDataSet$BLOCKXCOORD <- NULL
FinalDataSet$BLOCKYCOORD <- NULL
FinalDataSet$ANC <- NULL
FinalDataSet$DISTRICT <- NULL
FinalDataSet$PSA <- NULL
FinalDataSet$BUSINESSIMPROVEMENTDISTRICT <- NULL
FinalDataSet$BLOCK_GROUP <-NULL
FinalDataSet$CENSUS_TRACT <- NULL
FinalDataSet$VOTING_PRECINCT <- NULL

FinalDataSet <-  na.omit(FinalDataSet)


#Next, a new column stating the duration of the crimes was added to the dataset.The FinalDataSet had the columns "StartDate and an "EndDate". A new column was created with the name "DurationInDays" which had the difference of "StartDate" and "EndDate". Later on, the column "DurationInDays" was formatted to numeric for easy manipulation.

#[Step 6] Calculating the duration of crime
FinalDataSet$DurationInDays <- FinalDataSet$EndDate - FinalDataSet$StartDate
class(FinalDataSet$DurationInDays)
FinalDataSet$DurationNumeric <- as.numeric(FinalDataSet$DurationInDays)
FinalDataSet$DurationInDays <- NULL
#[Step 7] creating a ggplot histogram for the duration of crimes
library(ggplot2)

g <- ggplot(FinalDataSet, aes(x = DurationNumeric)) + geom_histogram(aes(), binwidth = 1)
g + scale_x_continuous(breaks = seq(min(FinalDataSet$DurationNumeric),max(FinalDataSet$DurationNumeric), by = 1)) + xlab("Duration of crimes in days") + ylab("Number of crime occurences") +labs(title = "Duration of Crimes and their frequency")
g


#For the purpose of analysis only those crimes were considered which had a duration, i.e. EndDate - StartDate, less than or equal to 7(the length of a week). Their difference denotes the time-period during which the crime could have occured. In this analysis I am assuming that the Holiday's effect on crime rates, if any, would be on the weeks adjacent to the holiday. Hence, nothing conclusive can be said about the effect of Holiday's on crimes that had a duration greater than 7 days.

#[Step 8] sorting the StartDate of crime in ascending order
FinalDataSet <- FinalDataSet[order(FinalDataSet$StartDate),]
min(FinalDataSet$StartDate, na.rm = TRUE)
FinalDataSet <- FinalDataSet[order(FinalDataSet$DurationNumeric),]

#[Step 9] Detecting outliers
summary(FinalDataSet$DurationNumeric)
#the value of duration of crimes range from -1659 days to 328700 days with dates from the year 1800 to 2911
#indicating that there are a lot of incorrect data

#[Step 10] Removing the outliers and keeping only the relevant values and keeping only those crime details for which duration was less than 7 days
FinalDataSet <- FinalDataSet[FinalDataSet$DurationNumeric <= 7 & FinalDataSet$DurationNumeric >= 0 & FinalDataSet$EndDate > as.Date("2011-01-01"),]

#[Step 11] removing the missing values
FinalDataSet <- na.omit(FinalDataSet)
dim(FinalDataSet)
table(FinalDataSet$OFFENSE)

#[Step 12] A new column "WeekNo" was added which contained the weeks corresponding to the "StartDate" for each crime.
#grouping the days into weeks
FinalDataSet$WeekNo <- format(FinalDataSet$EndDate, "%U")
FinalDataSet$WeekNo <- as.numeric(FinalDataSet$WeekNo)
#class(FinalDataSet$WeekNo)

#The second dataset containing the infrmation about all the holidays relevent to Washington DC wass imported. Relevant data-cleaning operations were performed on it.  A new column "Holidayweek" was added indicating the week no of each holiday. This "Holidayweek" would be used for analysing both the dataset.

#[Step 13] importing the holidays dataset
Holidays <- read.csv(file.choose())
#class(Holidays$Date)
Holidays$Date <- as.Date(Holidays$Date)
Holidays$WeekNo <- format(Holidays$Date, "%U")


#[Step 14] converting HolidaysWeek to numeric
Holidays$WeekNo <- as.numeric(Holidays$WeekNo)

#[Step 15] For a better insight into the effect of crimes the FinalDataSet wass categorized into several other subsets depending on the type of offenses(Robbery, Theft etc)

table(FinalDataSet$OFFENSE)
ArsonSet <- subset(FinalDataSet, FinalDataSet$OFFENSE == "ARSON")
AssaultSet <- subset(FinalDataSet, FinalDataSet$OFFENSE == "ASSAULT W/DANGEROUS WEAPON")
BurglarySet <- subset(FinalDataSet, FinalDataSet$OFFENSE == "BURGLARY")
HomicideSet <- subset(FinalDataSet, FinalDataSet$OFFENSE == "HOMICIDE")
MotorTheftSet <- subset(FinalDataSet, FinalDataSet$OFFENSE == "MOTOR VEHICLE THEFT")
RobberySet <- subset(FinalDataSet, FinalDataSet$OFFENSE == "ROBBERY")
SexAbuseSet <- subset(FinalDataSet, FinalDataSet$OFFENSE == "SEX ABUSE")
TheftSet <- subset(FinalDataSet, FinalDataSet$OFFENSE == "THEFT/OTHER")
TheftAutoSet <- subset(FinalDataSet, FinalDataSet$OFFENSE == "THEFT F/AUTO")

#Both the "FinalDataSet" and "Holidays" were combined on the basis of "WeekNo" which, for the "FinalDataSet" had the week in which the crime was committed and the week in which the holiday was observed from the "Holidays" dataset. Two sample t-test were performed on the various datasets. The first sample consisted of the crime rates during holidays an the second consisted of crime rates during regular days (excluding holidays).

##[Step 16] Creating a function for conducting t-tests and using the library dplyr
HolidayVsNonHoliday <- function(x, y = Holidays){
  library(dplyr)
  w <- semi_join(x, y, by = "WeekNo" )
  w <- as.data.frame(table(w$WeekNo))
  z <- anti_join(x, y, by = "WeekNo")
  z <- as.data.frame(table(z$WeekNo))
  t.test(w$Freq, z$Freq)
}

#[Step 17] conducting t-tests on subsets
HolidayVsNonHoliday(ArsonSet)
HolidayVsNonHoliday(FinalDataSet)
HolidayVsNonHoliday(AssaultSet)
HolidayVsNonHoliday(BurglarySet)
HolidayVsNonHoliday(HomicideSet)
HolidayVsNonHoliday(MotorTheftSet)
HolidayVsNonHoliday(RobberySet)
HolidayVsNonHoliday(SexAbuseSet)
HolidayVsNonHoliday(TheftSet)
HolidayVsNonHoliday(TheftAutoSet)

#From the above t-tests we were not able to reject the null hypothesis. We can only state that Holidays DO NOT have any conclusive effect on the crime rates. However, from the t-tests, it is apparent that the mean number of crimes is lower during the holidays, especially Sex-Abbuses,having a p-value of 0.09, which was near our alpha value of 0.05.We then created a ggplot so that we can percieve the effects graphically

#[Step 18] ggplot
library(ggplot2)
a <- ggplot(FinalDataSet, aes(WeekNo))
##problem with binwidth
c <- a + geom_freqpoly(aes(col = OFFENSE), lwd = 1, binwidth = 1) + geom_vline(Holidays, aes(X =Holidays$WeekNo), xintercept = Holidays$WeekNo)
c + scale_x_continuous(breaks = round(seq(min(FinalDataSet$WeekNo), max(FinalDataSet$WeekNo), by = 2))) + xlab("Week Number") + ylab("Crime Occurances") + labs(title = "Effect of holidays on crime rates") + theme_gray()

#From the graph it is somewhat apparent that there is some kind of relation between the crime occurences and he holiday's weeks(shown by the black vertical lines here). The crime rates either go up or down during these weeks.
#Moreover, the crimes are exceptionally low during the Christmas and New Year weeks(for the weeks 0,1,52,53). It could well be that, since, people are generally in their homes, the criminals may be less willing to break into someone's house or commit any crime on a whole.
#Hence we conducted a t-test specifically for the weeks 0,1,52,53.
#Null hypothesis - The weeks 0,1,52,53 have no effect on crime rates.
#Alternate Hypothesis - There is atleast some effect of these weeks on the crime rates.

#[Step 19] estimating the effect of Christmas and New Year holidays on crime rate
ChristmasCrimes <- subset(FinalDataSet, FinalDataSet$WeekNo == 0 | FinalDataSet$WeekNo == 1 | FinalDataSet$WeekNo == 53 | FinalDataSet$WeekNo == 52)
#View(ChristmasCrimes)
ChristmasCrimes <- as.data.frame(table(ChristmasCrimes$WeekNo))
OtherCrimes <- subset(FinalDataSet, FinalDataSet$WeekNo != 0 & FinalDataSet$WeekNo != 1 & FinalDataSet$WeekNo != 53 & FinalDataSet$WeekNo != 52)
OtherCrimes <- as.data.frame(table(OtherCrimes$WeekNo))
t.test(ChristmasCrimes$Freq, OtherCrimes$Freq)


```

From the t-tests, again, we cannot say anything conclusive about the effect those weeks have on the crime rates. The p-value in this case, however, is fairly low (0.06) indicating that we may want to extend our observations to other years to make some conclusions.

Latest Updated: 14 December 2016. 01:49:19PM